--- This is an assignment of Quantium Virtual experience Program where i was tasked with -
    Creating and interpreting high level summaries of the data
    Finding outliers and removing these (if applicable)
    Checking data formats and correcting (if applicable)
--- I did this Exploratory data Analysis using R
install.packages("data.table")
install.packages("ggplot2")
install.packages("ggmosaic")
install.packages("readr")
#### Load required libraries
library(data.table)
library(ggplot2)
library(ggmosaic)
library(readr)
filePath <- "/cloud/project/quantium/"
transactionData <- fread(paste0(filePath,"QVI_transaction_data.csv"))
customerData <- fread(paste0(filePath,"QVI_purchase_behaviour.csv"))

```

```{r}
#### Examine transaction data

head(transactionData)
str(transactionData)
```

```{r Convert DATE to date format}
#### Convert DATE column to a date format
#### A quick search online tells us that CSV and Excel integer dates begin on 30 Dec 1899
transactionData$DATE <- as.Date(transactionData$DATE, origin = "1899-12-30")
str(transactionData)
```

```{r Summary of PROD_NAME}
#### Examine PROD_NAME
Generating a summary of the PROD_NAME column.
summary(transactionData$PROD_NAME)
transactionData[, .N, PROD_NAME]
```

```{r Further examine PROD_NAME}
#### Examining the words in PROD_NAME to see if there are any incorrect entries
#### such as products that are not chips
productWords <- data.table(unlist(strsplit(unique(transactionData[, PROD_NAME]), "
")))
setnames(productWords, 'words')
head(productWords)
```

```{r}

#### Removing digits

#### Removing special characters
productWords$words <- gsub("[^[:alpha:][:space:]]", "", productWords$words)
head(productWords)
#### Let's look at the most common words by counting the number of times a word appears and
#### sorting them by this frequency in order of highest to lowest frequency
productWords[, .N, words][order(N, decreasing = TRUE)]
```

```{r}
#### Remove salsa products
transactionData[, SALSA := grepl("salsa", tolower(PROD_NAME))]
transactionData <- transactionData[SALSA == FALSE, ][, SALSA := NULL]
```

```{r initial summary}
#### Summarise the data to check for nulls and possible outliers

summary(transactionData)

# Create a boxplot to check for outliers
boxplot(transactionData$TOT_SALES)
boxplot(transactionData$PROD_QTY)
```

```{r }
#### Filter the dataset to find the outlier
#Use a filter to examine the transactions in question.
transactionData_filtered <- transactionData[transactionData$PROD_QTY == 200, ]
```

```{r}
#### Let's see if the customer has had other transactions
#Use a filter to see what other transactions that customer made.
transactionData_filtered

transactionData_filtered1 <- transactionData[transactionData$LYLTY_CARD_NBR == 226000, ]
transactionData_filtered1
```

```{r}

# Remove rows where LYLTY_CARD_NBR is equal to 226000
transactionData <- transactionData[transactionData$LYLTY_CARD_NBR != 226000, ]
# View the modified data frame
str(transactionData)
```

```{r}
#### Count the number of transactions by date
#Create a summary of transaction count by date.
# Convert date column to a Date object
transactionData$DATE <- as.Date(transactionData$DATE, format = "%Y-%m-%d")

# Aggregate transaction count by date
transactions_by_date <- aggregate(x = list(count = transactionData$TXN_ID),
                                   by = list(date = transactionData$DATE),
                                   FUN = length)

# Print the summary
print(transactions_by_date)
#There’s only 364 rows, meaning only 364 dates which indicates a missing dat
```

```{r fig.align = "center"}
#### Create a sequence of dates and join this the count of transactions by date
# creating a column of dates that includes every day from 1 Jul 2018 to 30 Jun 2019, and join it onto the data to fill in the missing day.
# Generate a sequence of dates from 1 Jul 2018 to 30 Jun 2019
allDates <- data.table(seq(as.Date("2018/07/01"), as.Date("2019/06/30"), by =
"day"))
setnames(allDates, "DATE")
transactions_by_day <- merge(allDates, transactionData[, .N, by = DATE], all.x
= TRUE)
```

```{r}
head(transactions_by_day)
```

```{r fig.align = "center"}


#### Setting plot themes to format graphs
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
#### Plot transactions over time
ggplot(transactions_by_day, aes(x = DATE, y = N)) +
geom_line() +
labs(x = "Day", y = "Number of transactions", title = "Transactions over
time") +
scale_x_date(breaks = "1 month") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{r fig.align = "center"}
#### Filter to December and look at individual days

# Set the plot themes
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))

# Plot transactions over time, zoomed in to December
library(lubridate)

ggplot(transactions_by_day[month(DATE) == 12, ], aes(x = DATE, y = N)) +
geom_line() +
labs(x = "Day", y = "Number of transactions", title = "Transactions over
time") +
scale_x_date(breaks = "1 day") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

```{r Create pack size}
#### Pack size
#### We can work this out by taking the digits that are in PROD_NAME
transactionData[, PACK_SIZE := parse_number(PROD_NAME)]
#### Always check your output
#### Let's check if the pack sizes look sensible
transactionData[, .N, PACK_SIZE][order(PACK_SIZE)]
```
The largest size is 380g and the smallest size is 70g - seems sensible!
```{r }
#### Let's plot a histogram of PACK_SIZE since we know that it is a categorical variable and not a continuous variable even though it is numeric.

# Plot histogram of transactions by pack size
hist(transactionData[, PACK_SIZE])

```

```{r Create brand name}

#### Brands
# Over to you! Create a column which contains the brand of the product, by extracting it from the product name.
#### Checking brands
# Over to you! Check the results look reasonable.
### Brands
library(stringr)

transactionData$brand_name <- str_to_title(word(transactionData$PROD_NAME, 1))

# view the first 10 rows
head(transactionData$brand_name, 10)

unique(transactionData$brand_name)

```
Some of the brand names look like they are of the same brands - such as RED and
RRD, which are both Red Rock Deli chips. Let's combine these together.
```{r Clean brand names}
#### Clean brand names
transactionData[brand_name == "Red", brand_name := "Rrd"]
transactionData[brand_name == "Snbts", brand_name := "Sunbites"]
transactionData[brand_name == "Infzns", brand_name := "Infuzions"]
transactionData[brand_name == "Ww", brand_name := "Woolworths"]
transactionData[brand_name == "Smith", brand_name := "Smiths"]
transactionData[brand_name == "Ncc", brand_name := "Natural"]
transactionData[brand_name == "Dorito", brand_name := "Doritos"]
transactionData[brand_name == "Grain", brand_name := "Grnwves"]

# Over to you! Check the results look reasonable.
unique(transactionData$brand_name)
```

```{r 1 Exploratory data analysis}
#### Examining customer data
#Doing some basic summaries of the dataset, including distributions of any key columns.
summary(customerData)
tibble(customerData)
# Group the data by words and count the frequency of occurrence
word_freq <- customerData %>%
  group_by(LIFESTAGE) %>%
  summarize(count = n())
# Sort the data in descending order of frequency
word_freq <- word_freq %>%
  arrange(desc(count))
# View the top 10 most frequent words
head(word_freq, 10)

# Group the data by words and count the frequency of occurrence
word_freq <- customerData %>%
  group_by(PREMIUM_CUSTOMER) %>%
  summarize(count = n())
# Sort the data in descending order of frequency
word_freq <- word_freq %>%
  arrange(desc(count))
# View the top 10 most frequent words
head(word_freq, 10)

```
```{r }
#### Merge transaction data to customer data
data <- merge(transactionData, customerData, all.x = TRUE)
```

```{r Check for missing customer details}

# Find transactions without a matched customer
unmatched_transactions <- anti_join(transactionData, customerData, by = "LYLTY_CARD_NBR")

# Check if there are any unmatched transactions
if (nrow(unmatched_transactions) == 0) {
  cat("All transactions have a matched customer.\n")
} else {
  cat(paste0(nrow(unmatched_transactions), " transactions did not have a matched customer.\n"))
}


```
Great, there are no nulls! So all our customers in the transaction data has been
accounted for in the customer dataset.

```{r Code to save dataset as a csv}
fwrite(data, paste0(filePath,"QVI_data.csv"))
```
Data exploration is now complete!
## Data analysis on customer segments

```{r fig.width = 10, fig.align = "center"}
#### Total sales by LIFESTAGE and PREMIUM_CUSTOMER
#Calculating the summary of sales by those dimensions and create a plot.
sales_summary <- data %>% 
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>% 
  summarise(total_sales = sum(TOT_SALES)) %>%
  arrange(desc(total_sales))

# View the summary
sales_summary
library(ggplot2)

# Create the plot
ggplot(sales_summary, aes(x = LIFESTAGE, y = total_sales, fill = PREMIUM_CUSTOMER)) + 
  geom_col(position = "dodge") + 
  labs(title = "Total Sales by Lifestage and Customer Type",
       x = "Lifestage",
       y = "Total Sales",
       fill = "Customer Type") + 
  theme(plot.title = element_text(hjust = 0.5))

```

```{r fig.width = 10, fig.align = "center"}
#### Number of customers by LIFESTAGE and PREMIUM_CUSTOMER
#Calculating the summary of number of customers by those dimensions and create a plot.
customer_count <- data %>%
  count(LIFESTAGE, PREMIUM_CUSTOMER, name = "customer_count") %>%
  arrange(desc(customer_count))

# View the results
customer_count
library(ggplot2)

# Create the plot
ggplot(customer_count, aes(x = LIFESTAGE, y = customer_count, fill = PREMIUM_CUSTOMER)) + 
  geom_col(position = "dodge") + 
  labs(title = "Customer Count by Lifestage and Customer Type",
       x = "Lifestage",
       y = "Customer Count",
       fill = "Customer Type") + 
  theme(plot.title = element_text(hjust = 0.5))

```

```{r fig.width = 10, fig.align = "center"}
#### Average number of units per customer by LIFESTAGE and PREMIUM_CUSTOMER
#  Calculating and plot the average number of units per customer by those two dimensions.
prod_qty_avg <- data %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarize(avg_qty = mean(PROD_QTY))
# View the summary
prod_qty_avg

library(ggplot2)

# Create the plot
ggplot(prod_qty_avg, aes(x = LIFESTAGE, y = avg_qty, fill = PREMIUM_CUSTOMER)) +
  geom_col(position = "dodge") +
  labs(x = "LIFESTAGE", y = "Average Quantity per Customer",
       title = "Average Number of Units per Customer by LIFESTAGE and PREMIUM_CUSTOMER") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```


```{r fig.width = 10, fig.align = "center"}
#### Average price per unit by LIFESTAGE and PREMIUM_CUSTOMER
#Calculating and plot the average price per unit sold (average sale price) by those two customer dimensions.
price_per_unit <- data %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarize(avg_price_per_unit = sum(TOT_SALES) / sum(PROD_QTY))

# Plot average price per unit sold
ggplot(price_per_unit, aes(x = LIFESTAGE, y = avg_price_per_unit, fill = PREMIUM_CUSTOMER)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Customer Lifestage", y = "Average price per unit sold", title = "Average price per unit sold by customer dimension") +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{r}
#### Perform an independent t-test between mainstream vs premium and budget midage and
#### young singles and couples


# Filter the relevant data

t_test_data <- price_per_unit %>%
  filter((LIFESTAGE %in% c("MIDAGE SINGLES/COUPLES", "YOUNG SINGLES/COUPLES")) &
           (PREMIUM_CUSTOMER %in% c("Budget", "Premium")) |
           (LIFESTAGE %in% c("MIDAGE SINGLES/COUPLES", "YOUNG SINGLES/COUPLES")) &
           (PREMIUM_CUSTOMER == "Mainstream"))

group1 <- t_test_data %>%
  filter(LIFESTAGE %in% c("MIDAGE SINGLES/COUPLES", "YOUNG SINGLES/COUPLES") &
         PREMIUM_CUSTOMER == "Mainstream") %>%
  pull(avg_price_per_unit)

group2 <- t_test_data %>%
  filter(LIFESTAGE %in% c("MIDAGE SINGLES/COUPLES", "YOUNG SINGLES/COUPLES") &
         PREMIUM_CUSTOMER %in% c("Budget", "Premium")) %>%
  pull(avg_price_per_unit)

# Perform independent t-test
t_test_result <- t.test(group1, group2)

# View the test results
t_test_result
```

```{r fig.align = "center"}
#### Deep dive into Mainstream, young singles/couples
segment1 <- data[LIFESTAGE == "YOUNG SINGLES/COUPLES" & PREMIUM_CUSTOMER ==
"Mainstream",]
other <- data[!(LIFESTAGE == "YOUNG SINGLES/COUPLES" & PREMIUM_CUSTOMER ==
"Mainstream"),]

#### Brand affinity compared to the rest of the population
quantity_segment1 <- segment1[, sum(PROD_QTY)]
quantity_other <= other[, sum(PROD_QTY)]
quantity_segment1_by_brand <- segment1[, .(targetSegment =
sum(PROD_QTY)/quantity_segment1), by = BRAND]
quantity_other_by_brand <- other[, .(other = sum(PROD_QTY)/quantity_other), by
= BRAND]
brand_proportions <- merge(quantity_segment1_by_brand,
quantity_other_by_brand)[, affinityToBrand := targetSegment/other]
brand_proportions[order(-affinityToBrand)]
```

```{r fig.align = "center"}
#### Preferred pack size compared to the rest of the population

quantity_segment1_by_pack <- segment1[, .(targetSegment =
sum(PROD_QTY)/quantity_segment1), by = PACK_SIZE]
quantity_other_by_pack <- other[, .(other = sum(PROD_QTY)/quantity_other), by =
PACK_SIZE]
pack_proportions <- merge(quantity_segment1_by_pack, quantity_other_by_pack)[,
affinityToPack := targetSegment/other]
pack_proportions[order(-affinityToPack)]
```

```{r}
data[PACK_SIZE == 270, unique(PROD_NAME)]
```
